Reagents: expressing and composing patterns of fine-grained concurrency

### 2011-10-11 paper/paper.tex
some outlining for the paper

@@

\authorinfo{}{}{}
\title{Reagents}
\maketitle

\begin{abstract}
\end{abstract}

@@

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathpartir}

\newcommand{\ra}{\rightarrow}
\newcommand{\gives}{\vdash}
\newcommand{\proves}{\vdash}
\newcommand{\GA}{\ |\ }
\newcommand{\kw}[1]{\textsf{#1}}
\newcommand{\ty}[1]{\textrm{#1}\ }
\newcommand{\bind}{\mbox{\ \scriptsize$>\!\!>\!=$\ }}
\newcommand{\letrec}[2]{\kw{letrec}\ #1\ \kw{in}\ #2}
\newcommand{\LET}[4]{\kw{let}_{#1}\ #2 = #3\ \kw{in}\ #4}

@@

\authorinfo{}{}{}
% yikes!
\title{Reagents: expressing and composing patterns of fine-grained concurrency}
\maketitle

\begin{abstract}
blah blah
\end{abstract}

``Programs are what happens between cache misses.''

As with most ..., concurrency only exacerbates performance problems
with the memory hierarchy.

ahmdahl's law -- even if you parallelize the big chunks, the
fine-grained stuff will get you in the end

%\section{Reagent semantics}

%\begin{verbatim}

%\end{verbatim}

@@

# General notes

Imposes no cache-coherence overhead on isolated data structures

Can argue that transactional-event-style liveness properties are
important if you want general way of composing fine-grained concurrent
algorithms with message-passing

To what extent do reagents encompass STM?  who are we competing with?

multiparadim, expressive, fast -- all that's great, but sales pitch
needs to be oriented around a customer with specific needs.  any
reason not to use the pitch of extensible, composable j.u.c.?

# Contributions

  - make it easier to express fine-grained concurrent algorithms,
  including blocking operations etc.  make it possible to build them
  compositionally, which already happens in the literature but is
  largely unremarked upon.

  - we capture exponential backoff, elimination backoff, and other
  such patterns once and for all

  - can use flat combining as illustration of capturing very general
  concurrency abstraction

  - smooth blend of message-passing and shared-state concurrency
  isolation for shared state, coherent communication through messages,
  uniform blocking/synchronization

  - operational semantics

  - yields completely lock-free implementation of core CML

  - generalizes join calculus implementation, allowing dissolution
  even for existing channels

  - generalizes transactions events, but with much better
  implementation

# To do

  - benchmark against STM (possibly Akka Transactors?)

  - benchmark queues, including flat combining and buckets

# Outline

2.5 Title, abstract, intro, library overview 
1.5 Semantics 
2.5 Implementation/algorithms 
1.0 Performance results
1.5 Related work
1.0 Conclusion and bib

How much time should we spend covering scalable concurrency
background?  This is perhaps tied to strategies for introducing the
library constructs: the primitives can be shown by-need for examples.
Should consult good papers like Ryan's and Jesse's for examples here.

# Semantics

Need to nail down story for overlapping state access.  In particular,
when (if ever) is an error flagged?  Unfortunately, talking about heap
unsplittability is rather more complicated than splittability.
